\documentclass{nlctdoc}

\usepackage[utf8]{inputenc}
\ifpdf
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\usepackage[scaled=.88]{helvet}
\usepackage{courier}
\usepackage{xr-hyper}
\fi
\usepackage[colorlinks,pdfauthor={Nicola L.C. Talbot},
            pdftitle={The glossaries package: a guide for beginners}]{hyperref}

\title{The glossaries package: a guide for beginners}
\author{Nicola L.C. Talbot}
\date{2011-04-12}

\newcommand*{\igloskey}[2][newglossaryentry]{\icsopt{#1}{#2}}
\newcommand*{\gloskey}[2][newglossaryentry]{\csopt{#1}{#2}}

\ifpdf\externaldocument{glossaries-user}\fi

\begin{document}
\maketitle

\begin{abstract}
This document is a brief guide to the \styfmt{glossaries} package for
beginners who find the size of the main user manual daunting and, as
such, it only covers the basics. For more advanced topics, see the
\docref{main user manual}{glossaries-user}.
\end{abstract}

\tableofcontents

\section{What Do I Need?}
\label{requirements}

You need to have the \styfmt{glossaries} package installed. You also
need to have the following packages installed: \sty{ifthen},
\sty{xkeyval} (at least version 2.5f), \sty{xfor}, \sty{amsgen} and
\sty{etoolbox}.
You may also need \sty{longtable}, \sty{supertabular} and
\sty{array}, depending on what glossary style you want to use. All
these packages can be downloaded from CTAN if you don't have them
installed. See
\urlfootref{http://www.tex.ac.uk/cgi-bin/texfaq2html?label=installthings}{\qt{Installing
things on a (La)TeX System}} for further assistance.

In addition, you need an indexing application. The \sty{glossaries}
package is configured to use either \app{makeindex} or \app{xindy}.
\app{Makeindex} should be available on all \TeX\ systems. It's
possible you may not have \app{xindy} installed, but it can be
downloaded from the \urlfootref{http://www.xindy.org/}{xindy home
page}. MiKTeX users may find the
\urlfootref{http://www.latex-community.org/forum/viewtopic.php?f=51&t=5383}{Xindy for MiKTeX} thread on the LaTeX Community Forum helpful.

The \sty{glossaries} package provides a helper Perl script
called \app{makeglossaries} which will act as an interface to
\app{makeindex} or \app{xindy}. If you want to use
\app{makeglossaries} or \app{xindy}, you need to have Perl
installed. If you don't have Perl installed, you can still use
\styfmt{glossaries} but you can only use it with \app{makeindex}.

\section{Getting Started}
\label{start}

The most obvious things to do is to remember to load the 
\sty{glossaries} package:
\begin{verbatim}
\usepackage{glossaries}
\end{verbatim}
\begin{important}
Note that the \styfmt{glossaries} package must be loaded \emph{after}
the \sty{hyperref} package (contrary to the general advice that
\sty{hyperref} should be loaded last). The \sty{glossaries} package 
should also be loaded after \sty{html}, \sty{inputenc},
\sty{babel} and \sty{ngerman}.
\end{important}

If you want to use \app{xindy} rather than \app{makeindex} to sort the
glossaries, you need to specify this in the package options:
\begin{verbatim}
\usepackage[xindy]{glossaries}
\end{verbatim}
Other package options are described in 
\xrsectionref{sec:pkgopts}{glossaries-user}{Package Options}
in the main \sty{glossaries} user manual.

Next, you need to enable the creation of the various external
files that are required by the indexing application. This is done using:
\begin{definition}
\cs{makeglossaries}
\end{definition}

\section{Defining a Term or Symbol}
\label{defterm}

You must define terms or symbols before you can use them in the
document. This is best done in the preamble. Each term is defined
using:
\begin{definition}
\cs{newglossaryentry}\marg{label}\marg{settings}
\end{definition}
where \meta{label} is a unique label used to identify the term.
The second argument, \meta{settings}, is a key=value comma 
separated list that is used to set the required information for
the term. A full list of available keys can be found in 
\xrsectionref{sec:newglosentry}{glossaries-user}{Defining Glossary
Entries}
in the main \sty{glossaries} user manual. The principle keys are
\gloskey{name} and \gloskey{description}.

For example, to define the term \qt{electrolyte}:
\begin{verbatim}
\newglossaryentry{electrolyte}{name=electrolyte,
description={solution able to conduct electric current}}
\end{verbatim}
In the above example, the label and the name happen to be the same.
In the next example, the name contains a ligature but the label
doesn't:
\begin{verbatim}
\newglossaryentry{oesophagus}{name=\oe sophagus,
description={canal from mouth to stomach},
plural=\oe sophagi}
\end{verbatim}
In this case the label is \texttt{oesophagus}. Note that in this
case, the plural has to be specified using the \gloskey{plural} key
since it isn't formed by simply appending the letter \qt{s}. 

When you define terms, it is important to remember that they will
be sorted by an external application that doesn't interpret \LaTeX\
commands. \app{Xindy} is a little more \TeX-aware than \app{makeindex}, but
for this example they will both mis-sort \verb|\oe sophagus|.
Xindy will ignore the \verb*|\oe | part and will sort it according
to \texttt{sophagus}, so it will be placed with the terms starting
with an \qt{s}. \app{Makeindex}, on the other hand, will sort it according
to \verb*|\oe sophagus| which starts with a symbol (the backslash),
so it will end up in the symbol group. You therefore need to 
specify how to sort the term:
\begin{verbatim}
\newglossaryentry{oesophagus}{name=\oe sophagus,
description={canal from mouth to stomach},
plural=\oe sophagi,
sort=oesophagus}
\end{verbatim}
Note, however, that if you use the \sty{inputenc} package and
\app{xindy}, you can use the encoded character \texttt{œ} rather than
\cs{oe} and \app{xindy} will sort it correctly:
\begin{verbatim}
\newglossaryentry{oesophagus}{name=œsophagus,
description={canal from mouth to stomach},
plural=œsophagi}
\end{verbatim}

If the name starts with an accented character, the first character
must be grouped:
\begin{verbatim}
\newglossaryentry{elite}{name={\'e}lite,
description={select group of people},
sort=elite}
\end{verbatim}
or
\begin{verbatim}
\newglossaryentry{elite}{name={é}lite,
description={select group of people}}
\end{verbatim}

The name can also be a symbol, but you will also need to specify
the sort key. For example:
\begin{verbatim}
\newglossaryentry{pi}{name={\ensuremath{\pi}},
description={ratio of circumference of circle to its 
diameter},sort=pi}
\end{verbatim}
It's also possible to have a name and a symbol. For example:
\begin{verbatim}
\newglossaryentry{set}{name=set,
description={a collection of objects},
symbol={\ensuremath{\mathcal{S}}}}
\end{verbatim}
Note, however, that not all glossary styles display the symbol
in the glossary.

\section{Using a Term}
\label{usingterm}

Once you have defined a term, you can use it in the document. The
\sty{glossaries} package provides a number of commands that are
described in \xrsectionref{sec:glslink}{glossaries-user}{Links to
Glossary Entries} in the main \sty{glossaries} user manual.
Here, I shall just cover the main commands:
\begin{definition}
\cs{gls}\marg{label}
\end{definition}
This prints the term associated with \meta{label}. Using the
example in the previous section, \verb|\gls{oesophagus}|
will display \oe sophagus. If the \sty{hyperref} package has also
been loaded, the term will also be hyperlinked to the relevant
entry in the glossary.
\begin{definition}
\cs{glspl}\marg{label}
\end{definition}
This prints the plural of the term associated with \meta{label},
so \verb|\glspl{oesophagus}| will display \oe sophagi. Again, the
text will be a hyperlink to the relevant entry in the glossary, if
hyperlinks are defined.

There are also versions that convert the first character to upper
case if you need to start a sentence with a term:
\begin{definition}
\cs{Gls}\marg{label}
\end{definition}
and
\begin{definition}
\cs{Glspl}\marg{label}
\end{definition}
For example:
\begin{verbatim}
\Glspl{electrolyte} usually exist as solutions of salts, 
bases or acids.
\end{verbatim}

Note that these commands all have two optional arguments that aren't
discussed here, but they are described in
\xrsectionref{sec:glslink}{glossaries-user}{Links to Glossary
Entries} in the main \sty{glossaries} user manual.

\begin{important}
Don't use these commands in moving arguments, such
as those used by \cs{chapter}, \cs{section} and \cs{caption}.
\end{important}
You can, instead, use commands such as
\begin{definition}
\cs{glsentrytext}\marg{label}
\end{definition}
See \xrsectionref{sec:glsnolink}{glossaries-user}{Using Glossary
Terms Without Links} in the main \sty{glossaries} user manual.

\section{Acronyms}
\label{acronyms}

The \sty{glossaries} package also allows you to define acronyms.
By default, acronyms will be added to the main glossary, but if
you use the \pkgopt{acronym} package option, acronyms will be
placed in a separate list of acronyms:
\begin{verbatim}
\usepackage[acronym]{glossaries}
\end{verbatim}
There are other options that govern how acronyms appear. For further
details, see \xrsectionref{sec:acronyms}{glossaries-user}{Acronyms}
in the main \sty{glossaries} user manual.

Acronyms must be defined before use and are defined using:
\begin{definition}
\cs{newacronym}\marg{label}\marg{abbrv}\marg{full}
\end{definition}
where \meta{label} is the label identifying the acronym, \meta{abbrv}
is the abbreviated form and \meta{full} is the fully expanded text.
For example:
\begin{verbatim}
\newacronym{label}{svm}{support vector machine}
\end{verbatim}

Once the acronym is defined, it can be used in the same way as 
any other glossary term with commands such as \cs{gls} and
\cs{glspl}. On first use, the default will display \meta{full}
(\meta{abbrv}). On subsequent use, only the abbreviation will
appear. For example:
\begin{verbatim}
A \gls{svm} was implemented.
\end{verbatim}
If this is the first time this term has been used, the above would 
produce:
\begin{display}
A support vector machine (svm) was implemented.
\end{display}
otherwise it would produce:
\begin{display}
A svm was implemented.
\end{display}
You can reset an acronym so that the next use fully 
expands it again using:
\begin{definition}
\cs{glsreset}\marg{label}
\end{definition}
or you can reset all acronyms using
\begin{definition}
\cs{glsresetall}
\end{definition}
For further details on resetting and unsetting acronyms, see
\xrsectionref{sec:glsunset}{glossaries-user}{Unsetting and
Resetting Entry Flags}
in the main \sty{glossaries} user manual.

\section{Displaying the Glossary or List of Acronyms}
\label{printglossaries}

In order to display the sorted list of terms, you first need to
use
\begin{definition}
\cs{printglossaries}
\end{definition}
at the place where you want the glossary and list of acronyms to
appear.  Note that \cs{printglossaries} won't produce any text the
first time you create your document, as an external indexing
application needs to be used to sort and collate the entries.
Creating the completed document is therefore \emph{at least} a three
step process.

Suppose your document is called \texttt{myDoc.tex}, then you need
to:
\begin{enumerate}
\item Build/\LaTeX\ your document, either by clicking on an
appropriate button in your front-end/text editor or by typing the
following at a terminal/command prompt:
\begin{verbatim}
latex myDoc
\end{verbatim}

\item Run the indexing application on the external file(s) created
by the \sty{glossaries} package. If you have used the \pkgopt{xindy}
package option, you must use \app{xindy}, otherwise you must use 
\app{makeindex}. The \sty{glossaries} package comes with a convenient
Perl wrapper that calls the appropriate indexing application the
required number of times with the required options. If you have Perl
installed, all you need to do is type
\begin{verbatim}
makeglossaries myDoc
\end{verbatim}
in a terminal or command prompt. If you are using a front-end (such
as WinEdt or TeXnicCenter) and you want it to automatically run
makeglossaries when you build your document, you will need to
consult the front-end's documentation.

If you don't have Perl installed, you will need to call
\app{makeindex} or \app{xindy} explicitly. This is more complicated
and is described in
\xrsectionref{sec:makeglossaries}{glossaries-user}{Generating the
Associated Glossary Files} in the main \sty{glossaries} user manual.

\item Once you have successfully run the indexing application, you
need to build/\LaTeX\ your document again (see step~1).
\end{enumerate}
Note that sometimes you may have to repeat steps~2 and~3.

The appearance of the glossary depends on the glossary style in
use. The default style uses the \env{description} environment, 
but there are many other styles available. These are described
in \xrsectionref{sec:styles}{glossaries-user}{Glossary Styles}
in the main \sty{glossaries} user manual.
By default, a number or list of numbers will appear after each entry 
in the glossary. This indicates the page (or pages)
on which the term was used. This list can be suppressed using
the \pkgopt{nonumberlist} package option:
\begin{verbatim}
\usepackage[nonumberlist]{glossaries}
\end{verbatim}

Further information can be found in the main \sty{glossaries}
\docref{user manual}{glossaries-user}
and there is also an
\href{http://www.latex-community.org/index.php?option=com_content&view=article&id=263:glossaries-nomenclature-lists-of-symbols-and-acronyms&catid=55:latex-general&Itemid=114}{article on the glossaries package}
on the \urlfootref{http://www.latex-community.org/}{\LaTeX\ Community's}
Know How section.
\end{document}
